---
title: "TaylorSwift"
author: "Eric Allen"
date: "10/22/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)

```

## Loading the data 

Album, Sales, Year 

Need to summarize into total album sales

```{r TaylorData}

tay<- read.csv('/Users/erallen/JHUCourse/taylorswiftspotify.csv')
dfalbumsales<- read.csv('/Users/erallen/JHUCourse/albumsales.csv')
#  dfcontpop<- read.csv('/Users/erallen/JHUCourse/population-regions-with-projections.csv')

```

## Data Shaping

1. create genre/album data
2. Need to merge on genre by album
3. Need to summarize into total album sales by continent
4. Compute Z scores - Normalize album poplularity 

- Album, Continent, Sales, ContinentPop, PopNormedSales

Summarize based on continent

- Continent, Sales, ContinentPop, PopNormedSales


# 1. Genre data 

```{r Genre}

# Load the tidyverse package (includes tidyr for data reshaping)
library(tidyverse)

# Original data
album_data <- data.frame(
  Album = c("1989", "Evermore", "Fearless", "Folklore", "Lover", "Red", "Reputation", "Speak Now", "Taylor Swift"),
  Alternative = c(0, 17, 0, 17, 0, 0, 0, 0, 0),
  Country = c(0, 0, 19, 0, 0, 0, 0, 17, 15),
  Pop = c(16, 0, 0, 0, 18, 22, 15, 0, 0)
)

# Reshape the data to have Album and Genre fields, eliminating rows with 0 sales
reshaped_data <- album_data %>%
  gather(key = "Genre", value = "Songs", -Album) %>%
  filter(Songs > 0)  # Filter out rows where sales are 0

# Print the reshaped data
print(reshaped_data)


```

# 2. Merge Genre Data 

```{r Genre Merge}


# Merge genre information onto sales_data based on the Album column
merged_data <- merge(dfalbumsales, reshaped_data, by = "Album", all.x = TRUE)

# Print the merged data
print(merged_data)



```


# 3. Need to summarize into total album sales by continent

```{r sales by cont}

continent_sales <- merged_data %>%
  group_by(Album, Continent, Genre) %>%
  summarize(TotalSales = sum(Sales), .groups = 'drop')

# Print the summary data
print(continent_sales)

```


# 4. **Calculate Z-Scores:**

To determine if certain genres are more popular on certain continents while normalizing the sales data to remove the impact of overall album popularity, you can use a technique called z-score normalization. Z-score normalization, also known as standardization, transforms your data to have a mean of 0 and a standard deviation of 1. This allows you to compare the popularity of genres within each continent without the influence of varying overall album sales.

Here are the steps to do this in R:

First, calculate the mean and standard deviation of album sales for each continent. Then, use these values to calculate the z-scores for the sales data within each continent.


```{r z-score}

# Calculate mean and standard deviation of sales for each continent
continent_stats <- continent_sales %>%
  group_by(Continent) %>%
  summarize(mean_sales = mean(TotalSales), sd_sales = sd(TotalSales))

# Merge z-scores back into the original data frame
normalized_data <- continent_sales %>%
  left_join(continent_stats, by = "Continent") %>%
  mutate(z_score = (TotalSales - mean_sales) / sd_sales)
```

In this code:
- `mean_sales` and `sd_sales` are the mean and standard deviation of sales for each continent.
- `z_score` calculates the z-score for each album's sales within its continent.

### 2. **Analyze Normalized Data:**

Now, you can analyze the `z_score` column to compare genre popularity within continents while removing the impact of overall album sales.

```{r AnalyzeNormalizedData}
# Analyze normalized data
genre_continent_summary <- normalized_data %>%
  group_by(Genre, Continent) %>%
  summarize(mean_z_score = mean(z_score), .groups = 'drop')

# Print the summary
print(genre_continent_summary)
```

In this summary, `mean_z_score` represents the average z-score of album sales for each genre within each continent. Positive or negative z-scores indicate sales above or below the continent's mean, respectively, after adjusting for overall album popularity.

By analyzing `mean_z_score`, you can identify which genres are relatively more or less popular within specific continents, considering the impact of overall album sales.

```{r }

# Load necessary libraries
library(reshape2)

# Melt the data for heatmap
melted_data <- melt(genre_continent_summary, id.vars = c("Genre", "Continent"))

# Create a heatmap
ggplot(data = melted_data, aes(x = Genre, y = Continent, fill = value)) +
  geom_tile() +
  labs(title = "Z-Score Heatmap by Genre and Continent",
       x = "Genre",
       y = "Continent",
       fill = "Mean Z-Score") +
  theme_minimal() +
  scale_fill_viridis_c()

```

```{r world heatmap}

library(leaflet)
library(htmlwidgets)
library(maptools) - not available for this verion of R
library(rgdal) - not available for this version of R
library(rworldmap)

# Load world map data
world_map <- getMap()

# Merge z-score data with world map data
merged_data <- merge(world_map, your_data, by.x = "NAME", by.y = "Continent", all.x = TRUE)

# Create the world map heatmap
world_map_heatmap <- leaflet(merged_data) %>%
  setView(lng = 0, lat = 0, zoom = 2) %>%
  addTiles() %>%
  addPolygons(fillColor = ~colorQuantile("YlOrRd", merged_data$mean_z_score)(mean_z_score),
              fillOpacity = 0.8,
              color = "white",
              weight = 1,
              smoothFactor = 0.5,
              label = ~paste(Genre, ": ", mean_z_score)) %>%
  addLegend(pal = colorQuantile("YlOrRd", merged_data$mean_z_score),
            values = ~mean_z_score,
            title = "Mean Z-Score",
            opacity = 0.7)

# Display the map
world_map_heatmap

install.packages("tmap")
install.packages("rworldmap")

library(tmap)
library(rworldmap)

# Load world map data
world_map <- getMap()

# Merge z-score data with world map data
merged_data <- merge(world_map, melted_data, by.x = "NAME", by.y = "Continent", all.x = TRUE)

# Plot the world map heatmap
tm_shape(merged_data) +
  tm_fill(col = "mean_z_score", style = "cont", palette = "YlOrRd", title = "Mean Z-Score") +
  tm_borders() +
  tm_layout(title = "World Map Heatmap by Genre and Continent")


```